Using VariantSpark to learn population structure from genomic profiles

### 2015-10-26 doc/publications/Recomb15/recomb15.tex
added my_plain.bst -- modified to reduce bib authors
small changes

% This is LLNCS.DOC the documentation file of
% the LaTeX2e class from Springer-Verlag
% for Lecture Notes in Computer Science, version 2.4
\documentclass{llncs}
\usepackage{llncsdoc}
\usepackage{placeins}
%

% Variable definitions
\newcommand{\variantSpark}{{\sc VariantSpark}}
\newcommand{\kMeans}{\textit{k}-means}
\newcommand{\ARI}{adjusted Rand index}

@@

Determining a person's ethnicity has become important for medical applications, such as HLA allele genotyping from SNP
information~\cite{Zheng2014}. Such clinical application are severely hampered for patients where accurate population
association is unknown due to a diverse migrational background. Determining the population association from genomic
information may provide vital insights for treatment and diagnostic in such cases.

@@

Using \kMeans{} we observed a perfect separation for the super populations African (AFR), East Asian (EAS) and European
(EUR), however when adding the fourth super-population group, American (AMR), we observe that only a subset of the AMR
individual form an independent cluster while the majority falls into the EUR cluster. 
We quantified this by calculating the \ARI\ (ARI), which compares the predicted and annotated labels. \kMeans achieved an ARI=0.84,
with -1 being random and 1 a perfect clustering.  We argue that the suboptimal clustering is a result of using an
unsupervised method rather than an insufficient number of genomic markers~\cite{Patterson2006} and hence hypothesise
that a supervised machine learning method may be capable of separating these subgroups of low genetic divergence.
Indeed, bigdata genomics recently investigated the supervised method, deep learning, for clustering the 1000 genomes
data by ethnicity~\cite{Ferguson}.

@@

%\marginpar{TODO do we need to cite ~\cite{Bureau2005,Yoo2012,Qi2012}.}  
We also investigate more complex supervised
approaches, namely Random Forest (RF), as for example Locke {\it et al.}~\cite{Locke2015} showed in a recent meta-study
around genetic susceptibility of obesity that only five of the 97 genomic loci have an independent association signal
the rest seem to jointly predict the complex metabolic phenotype.  We therefore investigate whether ethnicity is a
similarly complex phenotype that requires higher order dependencies between features to be taken into
account. %, as done by RF and DL.

@@

@@

\bibliographystyle{plain}
\marginpar{TODO you need to find a way of shortening the Locket paper author list}
\bibliography{recomb}

>>>>

% This is LLNCS.DOC the documentation file of
% the LaTeX2e class from Springer-Verlag
% for Lecture Notes in Computer Science, version 2.4
\documentclass{llncs}
\usepackage{llncsdoc}
\usepackage{placeins}
\usepackage{hyperref}
%

% Variable definitions
\newcommand{\variantSpark}{{\sc VariantSpark}}
\newcommand{\kMeans}{\textit{k}-means }
\newcommand{\ARI}{adjusted Rand index}

@@

\marginpar{I don't understand this. Can't you just sak a patient about their ancestry?}
Determining a person's ethnicity has become important for medical applications, such as HLA allele genotyping from SNP
information~\cite{Zheng2014}. Such clinical application are severely hampered for patients where accurate population
association is unknown due to a diverse migrational background. Determining the population association from genomic
information may provide vital insights for treatment and diagnostic in such cases.

@@

Using \kMeans{} we observed a perfect separation for the super populations African (AFR), East Asian (EAS) and European
(EUR), however when adding the fourth super-population group, American (AMR), we observe that only a subset of the AMR
individual form an independent cluster while the majority falls into the EUR cluster.  We quantified this by calculating
the \ARI\ ( ARI), which compares the predicted and annotated labels. \kMeans achieved an ARI of 0.84, with -1 being
random and 1 a perfect clustering.  We argue that the suboptimal clustering is a result of using an unsupervised method
rather than an insufficient number of genomic markers~\cite{Patterson2006} and hence hypothesise that a supervised
machine learning method may be capable of separating these subgroups of low genetic divergence.  Indeed, bigdata
genomics recently investigated the supervised method, deep learning, for clustering the 1000 genomes data by
ethnicity~\cite{Ferguson}.  \marginpar{``bigdata genomics'' seems to be a domain name \url{http://bdgenomics.org/} --
  how should we cite this}

@@

%\marginpar{TODO do we need to cite ~\cite{Bureau2005,Yoo2012,Qi2012}.}  
We also investigate more complex supervised
approaches, namely Random Forest (RF), as for example Locke {\it et al.}~\cite{Locke2015} showed in a recent meta-study
around genetic susceptibility to obesity that only five of the 97 genomic loci have an independent association signal
the rest seem to jointly predict the complex metabolic phenotype.  We therefore investigate whether ethnicity is a
similarly complex phenotype that requires higher order dependencies between features to be taken into
account. %, as done by RF and DL.

@@

\marginpar{not sure what is meant here. RF, LR etc are not multivariate -- there is only one $y$ variable, obesity. They
are multi-variable.}

@@

\bibliographystyle{my_plain}
\marginpar{FIXED:  TODO you need to find a way of shortening the Locket paper author list}
\bibliography{recomb}
